%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Laboratory Report LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original header:
%
% This is a LaTeX version of the sample laboratory report
% from Virginia Tech's copyrighted 08-09 CHEM 1045/1046 lab manual.
% Reproduction of this one appendix section for academic purposes
% should fall under fair use.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------


\documentclass{article}

    \usepackage[a4paper,vmargin={30mm,30mm},hmargin={20mm,20mm}]{geometry}
	\usepackage{natbib}		


\title{Location-aware Real Time Strategy Games \\ Master Thesis \\
Winter Semester 2012 - Summer Semester 2013}
% Title

\author{Vadim Costache} % Author name

\begin{document}

\maketitle % Insert the title, author and date

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}



%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS
%----------------------------------------------------------------------------------------
   
\begin{verbatim}



\end{verbatim}
     
\tableofcontents

\newpage


\section{Introduction: Real Time Strategy games}

The purpose of this paper is to describe the development of a new type of
GPS-enabled mobile game, a Real Time Strategy / Shooter hybrid. The motivation
for this came up during the search for GPS-enabled multiplayer game concepts.

Ever since the first operating system for handheld devices, the spreading of
smartphones has intensifed every year.In 2012, the number of smartphones in the
world has reached one billion, according to Bloomberg\cite{bloomberg} \newline

The rapid expansion of mobile computing presents new challenges and
opportunities for both the user and the developer. The ease of use and the
presence of touchscreens, GPS receivers, gyroscopes, accelerometers and, since
recently, even barometers gives way to new approaches in developing
games.\newline

In this paper, we will define a number of computer game genres, classify some of
the most popular location-aware games and game platforms and present the
evolution of the concept and development of a game type prototype. This game of
a genre that has a long legacy among computer games, but is not yet known to the
location-aware mobile context. The new genre is that of `Real Time
Strategy`(RTS)\cite{rts} games. We will do that with a subgenre
of RTS, called Real Time Tactics(RTT)\cite{rttvsrts}. For reasons that will be
presented throughout this paper, the type of game can be considered an
RTT / Shooter hybrid. \newline

\subsection{What is Real Time Strategy?}

RTS games are defined as real-time (continuous time) competitive games, in which
several players fight against each other, either in a skirmish or team versus
team. The purpose is to defeat the opponents by taking real-time decisions on
managing resources and troops. The focus is, therefore, on three key elements:
\textbf{managing resource gathering}, \textbf{building a base to provide troops}
and \textbf{battle tactics}.\cite{rts}\newline

This project will focus on a subgenre of RTS, Real Time Tactics(RTT) - which,
instead of managing all three aspects of RTS, focuses on \textbf{battle
tactics}.\newline

Because it is essentially a simplified approach to RTS\cite{rttvsrts}, RTT can
provide a good proof of concept and at the same time simple and enjoyable
playability. The advantage of having a RTT GPS-enabled game versus RTS is that
it greatly reduces the play time, requires less skills and has the potential of
being less stressful than RTS.\newline

\subsection{Real Time Strateg versus Turn Based Strategy}

Turn Based Strategy games\cite{rtsvstbs}(such as chess and board
games, for example) allow the player to take his time and plan every move.
Implicitly, the duration of the game is greater. This type of game is not in the
scope of this project, yet it deserves mentioning, as there are many such games
for mobile devices. Some notable implementations are the ones of Scotland
Yard(Ravensburger GMBH), Catan(Catan GMBH) and Monopoly(Hasbro, Inc.).\newline

In particular, there are two implementations for Scotland Yard upon which a
comparison has been made on whether it is better to port a board game in a
real-time(continuous-time) or turn-based mobile GPS-based game\cite{rttvsrts2}.
These two implementations are Mobilis XHunt(turn based) and MisterX
Mobile(real-time). The comparison has been made on 10 aspects: Fun, Smooth
Progression, Dynamic Gameplay, Ease of play, Stressless Gameplay, Communication,
Strategy, Clear Rules, Low Risk, Education. The conclusion was that there was no
favorite between the two, but it has been concluded that these 10 aspects have
different weights for the player and that Fun, Smooth Progression and Dynamic
Gameplay have a higher individual weight to players than the
rest.\cite[p. 5]{rttvsrts2}\newline

Based on the knowledge gathered from the above-mentioned comparison, the aim of
this project will be to maximize the interactivity and dynamics of the RTS game.


\section{Location-Based Augmented Reality Games}

Ever since the first operating system for handheld devices, the spreading of
smartphones has intensified every year.\newline

The rapid expansion of mobile computing presents new challenges and
opportunities for both the user and the developer. The ease of use and the
presence of touchscreens, GPS receivers, gyroscopes, accelerometers and, since
recently, even barometers gives way to new approaches in developing
games.\newline

Although it is one of the oldest additions to the smartphone, the GPS-enabled
smartphone is still not ubiquitous. As of now, each company's flagship
and most of their mid-level smartphones are GPS-enabled. This makes way to the
propagation of GPS gaming. Although the concept is old, very few attempts have
been made in this direction and this branch of game development may be
considered to still be in one of its early stages of maturity.\newline

This paper is a proposal for extending the genre of mobile real-time multiplayer
location-based games.

\subsection{Location-based games}
Location-based games take advantage of the mobile devices' built-in receivers
for global positioning. They provide the user's location with an accuracy
ranging from a few to a couple dozen meters. Because the most mobile devices in
the world today rely on the Global Positioning System(only recently support for
GLONASS has been added to smartphones), we can use the term GPS-games.\newline

GPS-games came up long before this feature has become ubiquitous in mobile
phones and tablets. One of the first widespread GPS-games is Geocaching. It is
composed of two parts: 
\begin{enumerate}
  \item Placing physical caches at various locations that can be considered
  interesting or worth visiting and publishing their GPS positions (eg. on
  websites).
  \item Searching for various caches by using a GPS device.
\end{enumerate}

Along with the evolution of smartphones came that of the mobile games. GPS games
come in a lot of flavors, from GPS-based tours, adventure and investigation
games to various race games - single and multi player and massively multiplayer
online games.


\subsection{Types of GPS-based games }

There is a number of GPS-enabled games and game authoring tools that are
available for iOS and Android devices. The ones studied for this paper are :
ARIS, Tourality, Wherigo, conTAGion, Shadow Cities, SCVNGR, Please Stay Calm,
Parallel Mafia, Parallel Kingdom, Tripventure, Warfinger, Totem, Portal Hunt,
aMazing, Ingress, MobileWar, Mister X Mobile, Mobilis XHunt, Own This World,
MapAttack.\newline

For better understanding the classification done below, we will first define
each type of game:

\begin{enumerate}
  \item \textbf{Adventure/Investigation Game} - Game in which the player plays
  the role of a character in a story. The primordial characteristics of this
  genre is that it is focused on immersion in the story, puzzle solving and
  investigation, rather than on physical skills. Also, the tendency of this
  genre is towards single player experience, though occasionally multiplayer is
  also implemented (eg. the ARIS-based game 'Mentira').
  
  \item \textbf{Massively Multiplayer Online Game} - This type of game is
  designed to support large numbers of players (even in the number of millions
  in some cases) that play and interact in a persistent virtual world. This type
  of game allows both cooperative and competitive gameplay and is exclusively
  based on multiplayer. Subgenres include MMO Role Playing Games and
  MMO Shooters.
  
  \item \textbf{Casual Game} - Analogous to the MMO, the Casual Game is targeted
  at mass at a mass audience and can incorporate any type of game type. The
  particularity of this genre is that it aims at having simple rules, simple
  gameplay and requiring no specialized skills. 
  
  \item \textbf{Racing Game} - It's a genre defining a broad range of games. In
  the case of computer games, it describes mostly motorized vehicle racing. In
  the mobile context, it mostly describes racing on foot against time or through
  a number of checkpoints.
  
  \item \textbf{Shooter Game} - This one's a subgenre of action games. It
  focuses on first or third person experience, speed, aiming and reaction time.
  Usually the weapon is ranged, although close-combat weapons are included in
  most games.
  
\end{enumerate}

We will now classify the games/platforms based on the genres they best fall in:

\begin{enumerate}
  \item \textbf{Adventure/Investigation Games}
  		\begin{enumerate}
  	  		\item ARIS
  	  		\item Wherigo
  	  		\item Tripventure
  	  		\item Tidy City	  
  		\end{enumerate}
  		
  \item \textbf{Massively Multiplayer Online Games}
  		\begin{enumerate}
  	  		\item Shadow Cities
  	  		\item Please Stay Calm
  	  		\item conTAGion
  	  		\item Parallel Mafia
  	  		\item Parallel Kingdom
  	  		\item Portal Hunt
  	  		\item Ingress
  		\end{enumerate}
  		
  \item \textbf{Casual Games}
  		\begin{enumerate}
  	  		\item SCVNGR
  	  		\item Warfinger  	  		
  	  		\item aMazing 	
  	  		\item Own This World
  	  		\item MapAttack	  	  		
  		\end{enumerate}
  		  		  		
  \item \textbf{Racing Games}
  		\begin{enumerate}
  	  		\item Tourality  	  		 	  		  	  		
  		\end{enumerate}
  		
  \item \textbf{Shooter Games}
  		\begin{enumerate}
  	  		\item MobileWar	  		  	  		
  		\end{enumerate}  
\end{enumerate}

A special category is represented by Mister X Mobile and Mobilis XHunt, which
both bring a board game (Scotland Yard) to the mobile environment. While the
former adapts the board game to real-time gameplay(placing it closer to the
'Multiplayer Racing Games', with elements from 'Real Time Strategy Games'), the
latter falls in the definition of 'Turn Based Strategy' games.

\section{Real Time Strategy Games}

This proposal is for the research and development of GPS-enabled Real Time
Strategy games. They are to be augmented reality games for single player or
multiplayer competitive 'free for all'/'skirmish' and 'team versus team'
games. This category of games offers opportunities to also enhance the
experience for all the previously existing types of GPS-enabled mobile games.

For this project, the proposed games are :
\begin{enumerate}	
	\item \textbf{Territory Takeover}
	\item \textbf{The War Game}
\end{enumerate}

1. The \textbf{Territory Takeover} game is a multiplayer, team versus team
competitive game. The players or game author define an area of play, which will
be divided into multiple divisions. Each division will be marked by a 'flag' (a
GPS marker). To capture the area division, a team must capture its flag. The
game ends when all flags have been captured and the winner is the team with most
captured flags. Each flag may be given a time that a player must spend next to
it in order to capture it. Once a flag (and implicitly the territory) is
captured, it remains so until the end of the game. The winner can be decided on
flag counting or, alternatively, each flag may receive a number of points,
according to the size of the territory marked by it and the difficulty of the
terrain. \newline

This game can be enhanced with the use of virtual tools or weapons. For the
purpose of this project, the following tools/weapons have been considered :
\begin{enumerate}
	
	\item The \textbf{Immobilizer} is an ability that can be used by each player to
	block an opponent from moving. The 'attacker' 'activates' the ability and a
	circle around him is drawn to show the range in which he can shoot. If an
	opponent enters the range area, the 'attacker' will select him on the map and
	shoot. The 'victim' will receive a notification that he is immobilized. A
	circle or rectangle will be defined around him and he will not be allowed to
	move outside of it for a given time, say 30 seconds or 1 minute. If he does, he
	gets disqualified and kicked out of the game. An alternate solution would be
	that the team loses points, for the case that this is the scoring methodology
	implemented.
	 
	\item \textbf{Demobilizer} is an ability that an immobilized player can use.
	For this project, it will only work on the person that uses the ability. The
	effect is that a person that is immobilized gets the waiting time halved.
	
\end{enumerate}
Both the abilities have a common cooldown timer. That means that if a player
immobilizes somebody and is immediately immoblized himself, he won't be able to
use the demobilizer because of the cooldown following the usage of the
immobilizer.

2. The War Game is inspired by Real Time Strategy Games and Airsoft. Two
teams are formed. An area of play is delimited on the map. Each player can
choose between a number of characters. For the purpose of this project, four
characters are proposed: Defender, Marine, Sniper and Heavy Trooper. Each of the
four characters has special abilities and characteristics :
\begin{enumerate}
	
	\item The \textbf{Defender} has the ability to generate shields for short
	periods of time. Members of the team can hide behind those shields for defence.
	The Defender may also act as a Medic and heal or revive members of the team. He
	has low health, long ability cooldowns and a sidearm with short range, small
	damage and fast cooldowns .
	
	\item The \textbf{Marine} has a weapon that can shoot a medium range with
	medium damage and fast cooldowns. He has medium health. 
	
	\item The \textbf{Sniper} has two weapons : the sniper rifle that can shoot at
	distant ranges and deal large damage to single targets and the sidearm, which
	is the same as the Defender's. His health is low, just like the Defender's. The
	sniper shot may penetrate the Defender's shield and cause reduced damage to one
	target.	
	
	\item The \textbf{Heavy Trooper} has three weapons : the bazooka, the sidearm
	and mines. The bazooka is a mid-range weapon with splash damage - it therefore
	can be fired against compact groups, such as the ones that might be hiding
	behind a shield. The bazooka cannot deal damage through the shield, but it may
	be shot next to it, causing damage from the side. The damage to each target
	varies from moderate to small, depending how far they are from the center of
	the 'projectile explosion'. The mines can be placed randomly on the map and
	their 'explosions' will not affect the members of the Heavy Trooper's team.
	Also they cannot be triggered by the members of his team. The damage dealt will
	be moderate, with splash damage, just like the bazooka projectile. The bazooka
	and the mines have long cooldowns, therefore the sidearm is added. The Heavy
	Trooper has high health.
	 
\end{enumerate}

This game has some advantages and drawbacks:\newline
\textbf{Advantages}: It does not require specialized gear and setting, nor does
it need long amounts of time to be played. It can be enjoyed with a bunch of
friends on a sunny weekend afternoon.\newline
\textbf{Drawbacks}: It highly depends on GPS accuracy. This issue may affect
gameplay. It also does not feel as 'real' as real-life games, nor PC
games.\newline


\section{Requirements}

The two above-mentioned games will use a common framework that will enable
multiplayer interaction for both 'free for all'/'skirmish' and 'team versus team'
approaches. They will require a server to centralize player information such as
GPS data and the virtual 'health' attribute. Because the games proposed are
fast-paced, they require quick response times from the server, client and and
the communication protocol between them.

\subsection{Client}
The following technologies have been considered for developing the frontend for
the application : Android, XCode(for Apple iOS) and multiplatoform APIs such as
PhoneGap/Cordova and Titanium. The multiplatform APIs offer the benefit of
the 'code once, deploy everywhere' philosophy, at the cost of
performance\cite{nativevscrossplatform}\cite{nativevscrossplatform2}\cite{nativevscrossplatform3}.
In this case, the approach of using a platform-agnostic framework is
disadvantageous. Therefore, the choice of native app development makes more
sense for this situation.\newline

The client will be developed with Android, due to its accessibility and
versatility and the fact that it's ubiquitous and open source.\newline

\subsection{Server}
The games that are to be implemented will actively rely on communication with a
centralized server. This implies a constant Internet connection on each mobile
device. Also, they require fast, low latency message exchanges. The server
should be able to handle this for players in the numbers of a few
dozens.\newline

The server will be written using NodeJS\cite{nodejs}. NodeJS is a javascript
framework based on the Google V8 Engine that is designed for web server
functionality. Yet, due to the fact that it is a framework and not a standalone
server, one can take advantage of the 'Event Loop' functionality to write
various other types of servers. This particular feat is beneficial to the
purpose of this project. It is also faster than Apache for general purpose, fast
message exchange scenarios\cite{nodejsvsapache}.

\subsection{Communication Protocols}
During a game, active communication must be performed between the mobile devices
and the server in order to multicast all player positions and intentions to all
players. There are two aspects to this approach:

\begin{enumerate}
	\item The game is created on the server, in a so-called `Lobby` which all the
	players join.
	\item The game is created on the server and once the game starts, all players send
	their positions to the server. The server is then responsible to multicasting
	the positions to all the players. 
\end{enumerate}

This is, of course, a subset of the tasks that must be performed by the server.
There will be other information that must be exchanged periodically, such as the
player's health and quantities of various other attributes that will be added
during development and/or proposed during trying out the game. In the general
idea of communication protocol usage, the list of choices has been narrowed down
to three:

\begin{enumerate}
	\item WebRTC - A protocol that is to be part of the HTML 5 standard. It will be
	based on the RTP(Reliable Transport Protocol), which is the base for VoIP
	protocols and is itself based on UDP. It promises to be a very fast standard
	protocol, appropriate for audio and video streaming and massive multiplayer
	games.\cite{webrtc} It is still in draft format and there are no official
	implementations for it. Implementing the protocol itself is outside the scope
	of this project.
	\item WebSockets - A protocol that is part of the HTML 5 standard. It is a low
	latency TCP-based protocol that promisses to replace http in several types of
	web applications.\cite{websockets}
	\item RTP - The protocol on which VoIP and WebRTC are based. It is based on UDP
	and it is designed for real-time streaming of data.\cite{rtp}
\end{enumerate}

From these three protocols, a choice will be made between WebSockets and RTP.
WebRTC is to be left as an option until its official release and implementation
for both NodeJS and Android.

\subsection{Game Elements}

Both games share a number of common features:
\begin{enumerate}
  \item a number of players that are visible on the map
  \item a number of teams, represented by different colors(in the case of
  free-for-all(skirmish) play, the number of teams equals the number of players)  
  \item a number of items that each player can use
  \item item properties, such as range and cooldown times
  \item player health
\end{enumerate} 

The 'War Game' also introduces a number of characters to the game. Each
character type has a different health value and different tools, with specific
properties.\newline

The server must provide a 'game lobby', to which the players connect and prepare
for starting the game. This includes character choice for the 'War Game' and a
'Ready' button. When all players declare themselves to be 'ready', a countdown
starts and then the game begins.\newline

\section{Implementation}

Implementation will mean developing a server and a client application from
scratch, covering all the functionality needed for the games to work according
to the description in section `Real Time Strategy Games`.

\subsection{Schedule}
This project, consisting of one server and one client application, is to be
developed in four steps: 
\begin{enumerate}
  \item \textbf{Development} - During the first iteration of development, the
  most basic features of the game are to be implemented: basic server
  functionality that would allow the game to work, basic client functionality
  and the 'War Game' without all features.(2 months)
  \item \textbf{Testing and Evaluation} - During this phase, the game and its
  functionality will be live-tested for feasibility and quality. New ideas will
  be sought and documented. Most importantly, player feedback will be
  gathered.(1 month)
  \item \textbf{Development} - During the second iteration of development, the
  'War Game' will be completed and, using its framework, the 'Territory
  Takeover' game will be implemented. Bugs will be removed and tweaks will be
  made to the framework and the game concepts to match the player feedback.(2
  months)
  \item \textbf{Evaluation and Completion} -  During the second evaluation
  phase, both games will be tested for playability, player feedback will be
  gathered and the Dissertation Paper will be completed.(1 month)
\end{enumerate}


\subsection{Authoring Tool}
As it is not the focus of this project, the Authoring Tool will comprise the
most basic elements necessary for setting up the game(such as defining obstacles
and probably the border of the gameplay area). Future work might include
modifying and/or adding new character attributes, multiple obstacle types and
probably extending the RTT games to full-fledged RTS.\newline

\section{Documentation on the Games}
Extensive documentation has been found from research done on education-oriented
GPS-games.\cite{pbarg1}\cite{pbarg2}\cite{pbarg3}\cite{pbarg4}\cite{pbarg5}\cite{pbarg6},
describing concepts and approaches in developing platforms and games to this
end\cite{pbarg3}, porting them to different locations\cite{pbarg4}, comparing
them and discussing their functionality\cite{pbarg6}\cite{pbarg1}\cite{pbarg2}
or discussing their effect and usefulness\cite{pbarg1}. Unfortunately for the
direction of this particular project, this documentation focuses exclusively on
GPS-enabled Adventure Games.\newline

The lack of documentation for the other games, except their websites has left
trying them out one by one as the only option to understand their components and
functionality. Additional information was retrieved from video descriptions,
examples and reviews of the games.\newline

During the preparation of this proposal paper no games or documentation have
been found on GPS-enabled Real Time Strategy games. From the searches performed,
no location-aware mobile games have been found to fall in the genre of RTS.
However, one has been found in the genre of `Shooter Games` - MobileWar.\newline

The purpose of this project is to create a framework on which the GPS-enabled
Real Time Strategy games 'Territory Takeover' and 'The War Game' will be
implemented as proof of concept. This framework is to be extended during future
work. The names of the games themselves are chosen only by their descriptive
nature and are prone to change during the steps of development and evaluation.

\section{Application of the Games}
The two games are proposed with group teambuilding and recreation in mind. Both
games require team strategy and cooperation. Territory Takeover allows for both
team versus team and free for all gameplay, allowing for both small and large
groups to play. The War Game is to be a fast paced game spanning a time interval
in the range of a few tens of minutes. Although it is a RTS Game and not a
Shooter Game, the 'War Game' can be seen as a more casual approach to Airsoft.
Where it lacks the realism of Airsoft or the immersion of classical computer
Real Time Strategy games, it gains in the intensity of real-life experience and
teamwork, without requiring specialized equipment(Airsoft) or highly developed
skills(computer Real Time Strategy games).


\section{The development of the game}

This section will follow the development of the application. The backend,
frontend and mechanics of the game will be analyzed and presented in their
evolution. \newline

The game is composed of three parts : \newline
\begin{enumerate}
  \item \textbf{Mobile App Frontend} - A set of menus for setting up the game
  and a GUI on top of Google Maps for the game itself.
  \item \textbf{Mobile App Backend} - The core of the app. This includes the
  modules for communicating with the server, for in-game protocols and
  mechanics.
  \item \textbf{Server} - All players will communicate with each other through a
  central server that can only host one game at a time.  
\end{enumerate}


\subsection{Initial plan}

The original plan was to develop and Android and iOS app, using a development
framework such as PhoneGap. This would assume Javascript / JQuery programming
for the UI and application backend. The communication was planned to be
established through Websockets. Therefore, for optimum performance, NodeJS was
to be used on the server side. For the location presentation, the Google Maps
API was to be used.

\subsection{Unified vs. Native Frameworks}

Once an idea of how the game should look like was crystalized, I started looking
for unified frameworks for cross-platform mobile development. The first choice
was PhoneGap, which I have used in the Agile Lab in the previous summer and was
a bit familiar. The other frameworks that I considered were Titanium SDK, Sencha
and Corona. The discussion on which one is better can be resumed to the
conclusion that none of them is appropriate for the development of such a game.
And here are the reasons : 
\begin{enumerate}
  \item The game is fast-paced and will require a fast backend as well as a
  frontend that is as fast as possible. The cross-platform frameworks
  essentially interface only some of the native functionality and display the
  app through a webview - requiring Javascript or a Javascript-based framework
  for creating the interfaces. I have tested some these frameworks on top of
  Phonegap, during the Agile Lab. The purpose was then to find the fastest
  solution for static menus and slow-paced UIs. Even then, the outcome has been
  disappointing. Qooxdoo, ExtJs and JQuery were tried out. Qooxdoo
  performs faster than the other two, but makes it very slow and tedious to
  develop a complex UI. Still, it is slow for the purpose and feels nonnative. 
  After reading a few articles that compare Phonegap, Titanium, Sencha and
  Corona, I have concluded that with or without various advantages and
  disadvantages between them, they are all similar in performance - and
  therefore too slow for the development of this game.
  
  \item Being a game prototype, all the details on which features of the phone
  or of the operating system will be used were unclear. I found it unwise to
  develop on a platform that only permits a number of features that are common
  to all platforms.
  
  \item Developing native code can prove to be easier, as for example the
  Android development community is much larger than that of Phonegap developers.
  The support is both more intensive and extensive for native platforms.
  
\end{enumerate}

From this point on, I gave up the unified frameworks and had to decide between
iOS and Android development. This was an easy choice : Android develpment is
free, while iOS development requires a paid developer account and XCode runing
on MacOS exclusively. Therefore, I chose Android. 


\subsection{Communication Protocols}

Initially, I have planned to use Websockets for the client-server communication.
The reasoning behind it has two main arguments : 
\begin{enumerate}
  \item Websockets is an HTML5 protocol currently presented in draft by the
  IETF. This protocol provides reliable two-way communication between a server
  and a client and manages various complex issues or network features that come
  above TCP. This future standard is developed to replace HTTP and add
  functionality for technical aspects that HTTP was not covering. The reason for
  using Websockets for the client-server communication was that it promises to
  abstract a lot of protocol management issues, while offering speeds close to
  plain TCP. Also, for the game to run properly, UDP and its child protocols
  cannot be used - total reliability is required in the communication. 
  
  \item Websockets communication can be implemented in NodeJS, which is a
  framework well suited for fast-paced message exchanges and which has been
  proven more scalable than, for example, Apache Tomcat. 
\end{enumerate}

As I decided to develop the server in NodeJS, the first choice was to use the
Socket.IO Websocket plugin. For the client side, I chose Autobahn for Android.
The only alternative for Autobahn was not free. After developing a basic
Websockets client with Autobahn for Android, I realized that they do not
communicate. After a lot of searching, I found out that the protocol
draft version that Socket.IO is using is different than that used in Autobahn,
and unlike Autobahn, Socket.IO uses an HTTP handshake for establishing the
connection. The Websocket-Node and ws NodeJS libraries for Websocket
communication were tried out afterwards. Neither were compatible with Autobahn
for Android. Then, I have switched to Python and after some technical
difficulties both in the Autobahn libraries for Android and establishing the
communication between client and server, I decided to give up Websockets and
start off with pure TCP. Because writing code in two different languages feels
slower, I have switched to Java.

\subsection{Maps API}

The first version of the app used Google Maps API V1. I chose it because it has
more extensive developer support. Unfortunately, the Google Maps API and the
Android Support Library cannot work together at the same time, because they need
to subclass different Fragment classes.

\subsubsection{Map Overlays}



 






\section{ANNEX A - game and game platform websites}



\nocite{teamtags}
\nocite{gps1}



\textbf{ARIS Games}: http://arisgames.org/ \newline

\textbf{Tourality} - http://www.tourality.com/ \newline

\textbf{Wherigo} - http://www.wherigo.com/ \newline

\textbf{conTAGion} - http://www.2clams.com/ \newline

\textbf{Shadow Cities} - http://www.shadowcities.com/ \newline

\textbf{SCVNGR} - http://www.scvngr.com/ \newline

\textbf{Please Stay Calm} - http://pleasestaycalm.com/ \newline

\textbf{Parallel Mafia} - http://www.parallelmafia.com/ \newline

\textbf{Parallel Kingdom} - http://www.parallelkingdom.com/ \newline

\textbf{Tripventure} - http://www.tripventure.net/tripventure/ \newline

\textbf{Warfinger GPS} - http://www.warfingergps.com/ \newline

\textbf{Totem} - http://www.fit.fraunhofer.de/en/fb/cscw/projects/totem.html
\newline

\textbf{Portal Hunt} - http://www.totem-games.org/?q=portalhunt \newline

\textbf{aMazing} - http://www.totem-games.org/?q=aMazing \newline

\textbf{Ingress} - http://www.ingress.com/ \newline

\textbf{Mobile War} - http://www.mobilewar.org/index.php/en/ \newline

\textbf{Mister X Mobile} - http://qeevee.com/projects/misterx \newline

\textbf{Mobilis XHunt} - https://github.com/danielschuster/mobilis/wiki/Mobilis-XHunt
\newline

\textbf{Own This World} - http://www.ownthisworld.com/ \newline

\textbf{MapAttack} - http://mapattack.org/ \newline



\section{ANNEX B - game genre definitions}

% Definitions - Game Genres %

\textbf{Real Time Strategy Game}		
http://encyclopedia2.thefreedictionary.com/Real-time+strategy+game
\begin{quote}
A type of video game in which players exercise strategy along the way, typically
to conquer enemies and reach a final destination without being eradicated. For
example, to win, players decide which routes to take, what needs to be done and
how to do it.
\end{quote}

\textbf{Real Time Tactics Game}
	http://encyclopedia.thefreedictionary.com/Real-time+tactics
\begin{quote}
Real-time tactics or RTT is a subgenre of tactical wargames played in real-time
simulating the considerations and circumstances of operational warfare and
military tactics. It is differentiated from real-time strategy gameplay by the
lack of resource micromanagement and base or unit building, as well as the
greater importance of individual units and a focus on complex battlefield
tactics.
\end{quote}

\textbf{Massively Multiplayer Online Game}		
	http://encyclopedia.thefreedictionary.com/Massively+Multiplayer+Online
\begin{quote}
A massively multiplayer online game (also called MMO) is a multiplayer video
game which is capable of supporting hundreds or thousands of players
simultaneously. By necessity, they are played on the Internet, and feature at
least one persistent world.
\end{quote}
	
\textbf{Adventure Game}		
	http://encyclopedia.thefreedictionary.com/adventure+game
\begin{quote}
An adventure game is a computer-based game in which the player assumes the role
of protagonist in an interactive story driven by exploration and puzzle-solving
instead of physical challenge.[1] The genre's focus on story allows it to draw
heavily from other narrative-based media such as literature and film,
encompassing a wide variety of literary genres. Nearly all adventure games are
designed for a single player, since this emphasis on story and character makes
multi-player design difficult.
\end{quote}

\textbf{Casual Game}		
	http://encyclopedia.thefreedictionary.com/casual+game
\begin{quote}
A casual game is a video game or online game targeted at or used by a mass
audience of casual gamers. Casual games can have any type of gameplay, and fit
in any genre. They are typically distinguished by their simple rules and lack of
commitment required in contrast to more complex hardcore games.[1] They require
no long-term time commitment or special skills to play[\ldots]
\end{quote}

\textbf{Racing Game}
	http://encyclopedia.thefreedictionary.com/Racing+game
\begin{quote}
One of the more common uses of the term racing game is to describe a genre of
computer and video games. Racing games are either in the first or third person
perspective. They may be based on anything from real-world racing leagues to
entirely fantastical settings, and feature any type of land, air, or sea
vehicles. In general, they can be distributed along a spectrum anywhere between
hardcore simulations, and simpler arcade racing games.
\end{quote}
\textbf{Shooter Game}
	http://encyclopedia.thefreedictionary.com/Shooter+game
\begin{quote}
Shooter games are a subgenre of action game, which often test the player's speed
and reaction time. Because shooters make up the majority of action games, it is
a fairly wide subgenre. It includes many subgenres that have the commonality of
focusing "on the actions of the avatar using some sort of weapon. Usually this
weapon is a gun, or some other long-range weapon".A common resource found in
many shooter games is ammunition. Most commonly, the purpose of a shooter
game is to shoot opponents and proceed through missions without dying yourself.
\end{quote}
	
\textbf{Turn Based Strategy Game}
	http://encyclopedia.thefreedictionary.com/Turn+based+strategy
\begin{quote}
A turn-based strategy (TBS) game is a strategy game (usually some type of
wargame, especially a strategic-level wargame) where players take turns when
playing. This is distinguished from real time strategy where all players play
simultaneously.
\end{quote}

\textbf{Location Based Game (Location-enabled Game}
	http://encyclopedia.thefreedictionary.com/location+based+game
\begin{quote}
A location-based game (or location-enabled game) is one in which the game play
somehow evolves and progresses via a player's location. Thus, location-based
games almost always support some kind of localization technology, for example by
using satellite positioning like GPS. "Urban gaming" or "Street Games" are
typically multi-player location-based games played out on city streets and built
up urban environments.
\end{quote}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------


\section{Bibliography}
	\bibliographystyle{ieeetr}	
	\bibliography{bibcontent}

\end{document}



