
\section{Future work}


This project has served as a proof of concept, a demonstration of the
potential GPS-based Real Time Tactics games. It has proven to be enjoyable
and easy to play for everybody who tried it.\newline

During the development and testing of this application, requests and ideas have
come up for future versions of the game. They will be presented in
order:\newline
\begin{enumerate}
  \item Icons will replace the names of all weapons and the professions of the
  players. A logo will be added and the menus will have suitable backgrounds.
  The current backgrounds are to be replaced.
  
  \item A further step in the development of the game would be to add RPG
  elements to it: The game mechanics will be modified to add elements of
  realism. Weapons will receive an optimum range attribute: Shotguns will be
  most effective in close range, rifles at longer range. Damage will be
  calculated based on range versus optimum range. 'Missing' a shot will become a
  factor in the game. Skills to use some weapons more effectively or to receive
  less damage will be subject of improvement. Health will also be subject of
  improvement. Two play modes will be added: casual and professional. In
  professional mode, players will be able to take the game to a further level
  and become competitive. They will be able to improve the damage dealt with
  weapons, if they used them a lot. The accuracy with certain types of weapons
  will be subject of improvement. In casual mode, all characters will be on the
  same level - thus allowing people to play, regardless of their daily
  implication in the game.
  
  \item Another aspect to the game will be adding an incentive for travel and
  discovery. Each player will be able to get temporary or permanent items by
  exploring the surroundings of his neighborhood or travelling to remote
  destinations.   
  
  \item An MMO game will be developed, having this game and its variants as
  subgames. The concept is as follows: An application will be linked to a social
  network, such as Facebook. Players will be able to log into the application
  with their Facebook accounts. They will set up who can see them on the map and
  who they want to see themselves: only themselves, friends, friends of friends,
  etc. They will be able to interact with each other directly through the app
  and invite each other to play games, such as the one developed here. Details
  such as creating custom avatars will be provided through the application or
  through a website. A scoring system will be added and competitions will be
  held through the app - they will be advertised through Facebook events.
  Joining the event will ensure participation in the competition. Facebook
  statuses might be published by the app, if the player wants so: entering,
  winning or losing a game, retrieving items, various accomplishments can be
  subject of publishing as Facebook status updates.
  
  \item Quests will be ultimately added to various locations, if it is deemed
  reasonable. This is still subject to debate, as the quests will likely be
  location-dependent and the game experience will then vary greatly from
  country to country or even from city to city.
  
\end{enumerate}

Other various technical modifications will be made to the game: In the first
improvement iteration, Websockets will be implemented as communication protocol.
Also, creating a server that would host a larger number of games is a priority
(now the server can only host one game).




