
\section{Conclusion}

The purpose of this project has been one of exploring a gap within the
conceptual construct of location-based mobile games, that of locaiton-based Real
Time Strategy games. Because no other implementation exists at the moment this
thesis is written, this has been an exploration into the unknown, a search for a
viable combination of gameplay elements that would bring enjoyment to the
players and help kickstart the concept.\newline

The exploration of ideas was carried throughout all genres of location-based
games(from Adventure to Massively Multiplayer Online(MMO) games) and a bit
further, among sports that predate the mobile phone itself. It has been noticed
that, like video games(PC and console), most location-based games focus more on
the story and immersion, than on social interaction and physical exercise.
Although there is a significant number of location-based multiplayer games,
their gameplay experience is largely solo and sedentary, with social interaction
narrowed down to a limited number of in-game actions.\newline

''People With Guns'' has addressed these aspects of gameplay, by exploring
Milgram's reality-virtuality continuum starting from a set target for the
reality aspect - a game that would entice players to move and communicate
directly - and adding virtuality in a suitable amount that would add to the game
experience.\newline

As the field testing has proven, the goal has been achieved: All the game
testers have enjoyed it: a roughly fifty percent female - fifty percent male
population, ranging from the age of 18 up to 27 have enjoyed playing and
requested the opportunity to play again. This has set a foundation from
which further development, testing and exploration can be done.\newline

The application developed during this project has proven to be socially engaging
mobile game to a broad spectrum of population, with or without technical skills
or gaming background. As half of the population of game testers has been of
non-gamers and they found it easy to get started and enjoy the game, we can
safely say that ''People With Guns'' has been a success on this thread of
exploration, too. \newline

With ''People With Guns'' - now called ''Gun Run'' - a new niche has been
created in the genre of location-based games. That is of location-based Real Time
Strategy games.\newline

