\appendix{Game genre definitions}

% Definitions - Game Genres %

\textbf{Real Time Strategy Game}		
http://encyclopedia2.thefreedictionary.com/Real-time+strategy+game
\begin{quote}
A type of video game in which players exercise strategy along the way, typically
to conquer enemies and reach a final destination without being eradicated. For
example, to win, players decide which routes to take, what needs to be done and
how to do it.
\end{quote}

\textbf{Real Time Tactics Game}
	http://encyclopedia.thefreedictionary.com/Real-time+tactics
\begin{quote}
Real-time tactics or RTT is a subgenre of tactical wargames played in real-time
simulating the considerations and circumstances of operational warfare and
military tactics. It is differentiated from real-time strategy gameplay by the
lack of resource micromanagement and base or unit building, as well as the
greater importance of individual units and a focus on complex battlefield
tactics.
\end{quote}

\textbf{Massively Multiplayer Online Game}		
	http://encyclopedia.thefreedictionary.com/Massively+Multiplayer+Online
\begin{quote}
A massively multiplayer online game (also called MMO) is a multiplayer video
game which is capable of supporting hundreds or thousands of players
simultaneously. By necessity, they are played on the Internet, and feature at
least one persistent world.
\end{quote}
	
\textbf{Adventure Game}		
	http://encyclopedia.thefreedictionary.com/adventure+game
\begin{quote}
An adventure game is a computer-based game in which the player assumes the role
of protagonist in an interactive story driven by exploration and puzzle-solving
instead of physical challenge.[1] The genre's focus on story allows it to draw
heavily from other narrative-based media such as literature and film,
encompassing a wide variety of literary genres. Nearly all adventure games are
designed for a single player, since this emphasis on story and character makes
multi-player design difficult.
\end{quote}

\textbf{Casual Game}		
	http://encyclopedia.thefreedictionary.com/casual+game
\begin{quote}
A casual game is a video game or online game targeted at or used by a mass
audience of casual gamers. Casual games can have any type of gameplay, and fit
in any genre. They are typically distinguished by their simple rules and lack of
commitment required in contrast to more complex hardcore games.[1] They require
no long-term time commitment or special skills to play[\ldots]
\end{quote}

\textbf{Racing Game}
	http://encyclopedia.thefreedictionary.com/Racing+game
\begin{quote}
One of the more common uses of the term racing game is to describe a genre of
computer and video games. Racing games are either in the first or third person
perspective. They may be based on anything from real-world racing leagues to
entirely fantastical settings, and feature any type of land, air, or sea
vehicles. In general, they can be distributed along a spectrum anywhere between
hardcore simulations, and simpler arcade racing games.
\end{quote}
\textbf{Shooter Game}
	http://encyclopedia.thefreedictionary.com/Shooter+game
\begin{quote}
Shooter games are a subgenre of action game, which often test the player's speed
and reaction time. Because shooters make up the majority of action games, it is
a fairly wide subgenre. It includes many subgenres that have the commonality of
focusing "on the actions of the avatar using some sort of weapon. Usually this
weapon is a gun, or some other long-range weapon".A common resource found in
many shooter games is ammunition. Most commonly, the purpose of a shooter
game is to shoot opponents and proceed through missions without dying yourself.
\end{quote}
	
\textbf{Turn Based Strategy Game}
	http://encyclopedia.thefreedictionary.com/Turn+based+strategy
\begin{quote}
A turn-based strategy (TBS) game is a strategy game (usually some type of
wargame, especially a strategic-level wargame) where players take turns when
playing. This is distinguished from real time strategy where all players play
simultaneously.
\end{quote}

\textbf{Location Based Game (Location-enabled Game}
	http://encyclopedia.thefreedictionary.com/location+based+game
\begin{quote}
A location-based game (or location-enabled game) is one in which the game play
somehow evolves and progresses via a player's location. Thus, location-based
games almost always support some kind of localization technology, for example by
using satellite positioning like GPS. "Urban gaming" or "Street Games" are
typically multi-player location-based games played out on city streets and built
up urban environments.
\end{quote}