\section{ANNEX A - Game and game platform websites}



\nocite{teamtags}
\nocite{gps1}



\textbf{ARIS Games}: http://arisgames.org/\newline

ARIS Games is a framework for developing Adventure/Investigation games. Most
applications of these games have been within the education area.\newline

\textbf{Tourality} - http://www.tourality.com/\newline

Tourality is a mobile application that allows people to play a number of racing
games. Among them are single lap or multiple lap races, capture the flag races
where players have to run towards a given marker on the map. The first one to
reach the marker gets one point - and another marker is randomly generated
elsewhere. This last race game has represented an important source of
inspiration in the construction of the idea for the game developed on this
project.\newline

\textbf{Wherigo} - http://www.wherigo.com/\newline

Just like ARIS, Wherigo is a toolset for creating Adventure/Investigation games.
\newline

\textbf{conTAGion} - http://www.2clams.com/\newline

Contagion is a GPS-based zombie apocalypse Massively Multiplayer Online game in
which the players are either in the role of zombies or humans, evading or
attacking each other. The game can also be played in single player mode. A game
has been attempted, but due to a difficult and buggy interface, it was
not considered worthwhile.\newline

\textbf{Shadow Cities} - http://www.shadowcities.com/\newline
 
Shadow cities is an MMMORPG(Massively Multiplayer Online Role Playing Game) in
which the player is put in the role of a mage. The mechanics of the game tend to
be similar to those of classical MMORPGs, but they now involve the map and the
GPS. The player has to travel around physically in order to reach various areas
where levelling may be done.\newline

\textbf{SCVNGR} - http://www.scvngr.com/\newline

SCVNGR is a social game that uses Google Places for checking in, creating or
accepting various challenges already left by others at given places.\newline

\textbf{Please Stay Calm} - http://pleasestaycalm.com/ \newline

Please Stay Calm is a single player GPS-based zombie apocalypse RPG game in
which the player has to scavenge or fight zombies in various areas from the
map.\newline


\textbf{Parallel Mafia} - http://www.parallelmafia.com/ \newline

Parallel Mafia is a GPS-based MMORPG in which the player builds up his own
mafia, cooperates with or fights others for territory. To this end, various tools
are present in the game. The ultimate goal is capturing as much territory as
possible.\newline

\textbf{Parallel Kingdom} - http://www.parallelkingdom.com/ \newline

Paralle Kingdom is a conceptually identical game to Parallel Mafia - only that
it is set in the middle ages.\newline

\textbf{Tripventure} - http://www.tripventure.net/tripventure/ \newline
Tripventure is a platform on which several augmented reality adventure games are
created. The special thing about these games is that they don't only use the
GPS, but also the phone's camera - at given GPS positions, virtual characters
must be sought and interacted with by looking at the phone screen and finding
them with the camera.\newline


\textbf{Warfinger GPS} - http://www.warfingergps.com/ \newline
Warfinger GPS is a casual game in which two teams fight within a given area,
over an indefinite amounte of time. The interaction is rather turn based than
real-time. This has been the biggest influence in the construction of the idea
for this project - some of the virtual objects and some of the weapons presented
in the game trailer have been redesigned for this project's game
proposals.\newline

\textbf{Totem} - http://www.fit.fraunhofer.de/en/fb/cscw/projects/totem.html
\newline
Totem is a versatile platform for mobile game authoring. Several games have been
created using this platform. Among them are Tidy City, Portal Hunt and aMazing -
which later conceptually appeared in Ingress.\newline

\textbf{Portal Hunt} - http://www.totem-games.org/?q=portalhunt \newline

\textbf{aMazing} - http://www.totem-games.org/?q=aMazing \newline

\textbf{Ingress} - http://www.ingress.com/ \newline
Ingress is the much advertised MMORPG from Google. An invitation is required to
play the game - therefore more information than what was presented in the
trailer is not in the posession of the author of this paper:\newline
\url{http://www.youtube.com/watch?feature=player_embedded&v=92rYjlxqypM}\newline

\textbf{Mobile War} - http://www.mobilewar.org/index.php/en/ \newline
Mobile War is a GPS-based shooting game - the players use their phones to aim
and 'shoot' each other in a classical Shooter game manner. This game has
represented an important source of inspiration for this project.
Unfortunately, it could not be played due to server errors.\newline

\textbf{Mister X Mobile} - http://qeevee.com/projects/misterx \newline
Mister X Mobile is one of the two mobile implementations for Ravensburger's
Scotland Yard game. This one takes a real-time approach. The game mechanics and
the use of various tools for catching 'Mr. X' have been an inspiration for the
games proposed\newline

\textbf{Mobilis XHunt} - https://github.com/danielschuster/mobilis/wiki/Mobilis-XHunt
\newline
Mobilis XHunt is the second implementation of Ravensburger's Scotland Yard game.
This one takes a turn-based approach. This game has not directly influenced this
project, the analysis\cite{rttvsrts2} of real-time versus turn-based approaches in
GPS-based mobile games that compares Mister X Mobile and Mobilis XHunt has.\newline

\textbf{Own This World} - http://www.ownthisworld.com/ \newline
Own This World is a game in which the world map is split into a grid of
'territories'. People physically present in one of these 'territories' can fight
for conquest. For this end, they use resources to create virtual troops and
fight each other. The more territory is available to the player, the more
resources he will receive - and therefore the easier he will conquer further
territories. This game has been the biggest source of inspiration for the
proposal of the 'Territory Takeover' game that is to be added to this project as
future work.\newline

\textbf{MapAttack} - http://mapattack.org/ \newline
MapAttack is another territory capture game based on the Geoloqi platform.
Unfortunately, the website is currently down. It can be otherwise found on GitHub:
\url{https://github.com/geoloqi/MapAttack}\newline
