
\section{Future work}


This project has served as a proof of concept, a demonstration of the
potential GPS-based Real Time Tactics games. It has proven to be enjoyable
and easy to play for everybody who tried it.\newline

During the development and testing of this application, requests and ideas have
come up for future versions of the game. They will be presented in
order:\newline
\begin{enumerate}
  \item Icons will replace the names of all weapons and the professions of the
  players. A logo will be added and the menus will have suitable backgrounds.
  The current backgrounds are to be replaced.
  
  \item A further step in the development of the game would be to add RPG
  elements to it: The game mechanics will be modified to add elements of
  realism. Weapons will receive an optimum range attribute: Shotguns will be
  most effective in close range, rifles at longer range. Damage will be
  calculated based on range versus optimum range. 'Missing' a shot will become a
  factor in the game. Skills to use some weapons more effectively or to receive
  less damage will be subject of improvement. Health will also be subject of
  improvement. Two play modes will be added: casual and professional. In
  professional mode, players will be able to take the game to a further level
  and become competitive. They will be able to improve the damage dealt with
  weapons, if they used them a lot. The accuracy with certain types of weapons
  will be subject of improvement. In casual mode, all characters will be on the
  same level - thus allowing people to play, regardless of their daily
  implication in the game.
  
  \item Another aspect to the game will be adding an incentive for travel and
  discovery. Each player will be able to get temporary or permanent items by
  exploring the sorroundings of his neighborhood or travelling to remote
  destinations.   
  
  \item An MMO game will be developed, having this game and its variants as
  subgames. The concept is as follows: An application will be linked to a social
  network, such as Facebook. Players will be able to log into the application
  with their Facebook accounts. They will set up who can see them on the map and
  who they want to see themselves: only themselves, friends, friends of friends,
  etc. They will be able to interact with each other directly through the app
  and invite each other to play games, such as the one developed here. Details
  such as creating custom avatars will be provided through the application or
  through a website. A scoring system will be added and competitions will be
  held through the app - they will be advertised through Facebook events.
  Joining the event will ensure participation in the competition. Facebook
  statuses might be published by the app, if the player wants so: entering,
  winning or losing a game, retrieving items, various accomplishments can be
  subject of publishing as Facebook status updates.
  
  \item Quests will be ultimately added to various locations, if it is deemed
  reasonable. This is still subject to debate, as the quests will likely be
  location-dependent and the game experience will then vary greatly from
  country to country or even from city to city.
  
\end{enumerate}

Other various technical modifications will be made to the game: In the first
improvement iteration, Websockets will be implemented as communication protocol.
Also, creating a server that would host a larger number of games is a priority
(now the server can only host one game).


\section{Conclusion}

The purpose of this project has been to develop a socially engaging mobile game
to a broad spectrum of population. In translation, the task was to make a game
that would be entertaining and enjoyable for people of both sexes and all ages,
with or without technical skills or gaming background. \newline

As the field testing has proven, the goal has been achieved: All the game
testers have enjoyed it: a roughly fifty percent female - fifty percent male
population, ranging from the age of 18 up to 27 have enjoyed playing and
requested the opportunity to play again. Because of the small scale of the
tests though, a larger age palette could not be tried out within the setting of
this project.\newline

With ''People With Guns'' - now called ''Gun Run'' - a new niche has been
created in the genre of location-based games. That is of location-based Real Time
Strategy games.\newline




%
%For me, the author, this has been a journey into how an idea is mined for and
%developed, step by step, try after try, using scientific reasoning
%alongside trial and error. Through this project, I have learned Android
%programming, applied Agile methodologies, devised a personal reasoning for the
%cases in which Unit Testing is applicable. I have made server prototypes with
%Node.JS, Python, Java, Map prototypes with the Google Maps V1 and V2 APIs,
% tried out Websockets implementations for Node.JS, Python and Android, tested
% the usage
%of both Google GSON and Jackson libraries for JSON serialization and
%deserialization. I have evolved an idea into a mature application with
% optimized data and hardware resource usage, with its own messaging and
% configuration
%system. The end result is a fast-paced game with a dynamic UI that has been
%optimized for ease of use and a set of 'weapons' and 'characters' that fit
%various types of player personalities. From now on, development will rely on a
%well set foundation.\newline
%
%
%Concomitant with the scientific approach of the conceptualization and
%development of this mobile game subgenre, this has been a very subjective,
%personal experience. A lot of intuition and communication have played a large
%part of what is now the real time tactics game prototype \"People With Guns\" .
%A lot of choices have been made on feeling and personal perception. The
% people's opinions on the application, design aspects and even the name of the
% game itself
%have been put to question and have received very varied feedback. I have seen
%hands-on what changing a background image from black to white can cause in the
%player's attitudes. I have seen difference of opinion and perspective between
%sexes and ages. This research has transcended the defined task of developing a
%game type that has not been tried before. It has become a very personal
%experience and came with its hardships and teachings. Although the entire paper
%is expressed scientifically, I have tried to write it in a manner as personal
% as possible.\newline

%The usage of 'he' should be replaced by 'she/he' or 'he/she' and has only been
%used this way for the convenience of the author and no other reasons.
%